\chapter{Performance Analysis}
\label{ch:performance}

\section{Model Performance Summary}

Table~\ref{tab:full_comparison} provides a comprehensive comparison of all models
and scenarios evaluated in this project.

\begin{table}[H]
\centering
\caption{Full Model Comparison — Test Set}
\label{tab:full_comparison}
\begin{tabular}{llrrrr}
\toprule
\textbf{Model} & \textbf{Horizon} & \textbf{MAE} & \textbf{RMSE} & \textbf{R²} & \textbf{AUROC} \\
\midrule
Persistence (baseline) & t+1 & 5.30  & ---   & ---   & --- \\
Persistence (baseline) & t+3 & 8.97  & ---   & ---   & --- \\
Persistence (baseline) & t+7 & 10.93 & ---   & ---   & --- \\
\midrule
LightGBM (champion)    & t+1 & \textbf{5.12}  & \textbf{9.00}  & \textbf{0.841} & \textbf{0.992} \\
LightGBM (champion)    & t+3 & \textbf{6.77}  & \textbf{11.78} & \textbf{0.729} & \textbf{0.973} \\
LightGBM (champion)    & t+7 & \textbf{8.25}  & \textbf{14.36} & \textbf{0.600} & \textbf{0.946} \\
\midrule
XGBoost (secondary)    & t+1 & 5.31  & 8.95  & 0.843 & 0.992 \\
XGBoost (secondary)    & t+3 & 7.08  & 11.90 & 0.723 & 0.973 \\
XGBoost (secondary)    & t+7 & 8.90  & 15.32 & 0.544 & 0.933 \\
\midrule
LSTM (comparison)      & t+1 & 6.67  & 11.64 & 0.753 & 0.984 \\
LSTM (comparison)      & t+3 & 8.87  & 15.89 & 0.540 & 0.927 \\
LSTM (comparison)      & t+7 & 8.06  & 14.49 & 0.608 & 0.961 \\
\midrule
XGBoost (no FIRMS)     & t+1 & 6.52  & 11.21 & 0.754 & 0.983 \\
XGBoost (no FIRMS)     & t+3 & 7.75  & 12.93 & 0.674 & 0.970 \\
XGBoost (no FIRMS)     & t+7 & 8.80  & 15.08 & 0.558 & 0.946 \\
\bottomrule
\end{tabular}
\end{table}

\section{Inference Latency}

The FastAPI \texttt{/forecast} endpoint loads both the LightGBM champion and the
XGBoost strong-secondary at startup and serves predictions from the pre-computed
latest feature row. A 5-minute in-process TTL cache keyed on the feature-file
modification time absorbs duplicate demo-time requests without staling when a new
feature matrix lands. Typical response times measured locally:

\begin{itemize}
    \item \texttt{/health}: $<$ 5~ms
    \item \texttt{/forecast}: $<$ 30~ms cold, $<$ 2~ms cached (3 horizons scored in sequence)
    \item \texttt{/predict} (POST): $<$ 50~ms
    \item \texttt{/history}: $<$ 100~ms (reads 60-day CSV slice)
\end{itemize}

Gradient-boosted inference is CPU-bound and scales linearly with the number of
input rows. A t3.small EC2 instance (2 vCPU, 2~GB RAM) with a 2~GB swap file is
sufficient for the current single-station, single-user workload with co-located
FastAPI, MLflow, and Streamlit.

\section{Scalability Analysis}

The current architecture supports one station and a low request rate.
Scaling paths:

\begin{itemize}
    \item \textbf{Multi-station:} Add stations by extending the data ingestion scripts
          and training one model set per station. S3 partitioning by station code
          keeps the pipeline clean.
    \item \textbf{High request rate:} FastAPI supports multiple Uvicorn workers
          (\texttt{--workers 4}) without code changes. Upgrading to a t3.small
          or t3.medium, or placing an Application Load Balancer in front of an
          Auto Scaling Group of EC2 instances, handles traffic spikes.
    \item \textbf{Real-time inference:} Current pipeline is daily-batch. Moving to
          hourly inference requires only a cron change in \texttt{daily\_pipeline.yml}.
\end{itemize}

\section{Cost Analysis}

% TODO: Replace estimates below with actual AWS console numbers after deployment.
% Run: aws ce get-cost-and-usage --time-period Start=YYYY-MM-DD,End=YYYY-MM-DD

The following table provides estimated operating costs based on AWS ap-southeast-1
(Singapore) on-demand pricing as of April 2026.

\begin{table}[H]
\centering
\caption{Estimated Monthly Operating Cost (AWS ap-southeast-1)}
\label{tab:cost}
\begin{tabular}{lrrr}
\toprule
\textbf{Component} & \textbf{Unit Cost} & \textbf{Monthly} & \textbf{Notes} \\
\midrule
EC2 t3.small (on-demand) & \$0.0208/hr & \$15.00 & FastAPI + MLflow + Streamlit co-located (2~GB RAM) \\
S3 storage (5~GB)        & \$0.023/GB  & \$0.12  & raw + processed + models + results \\
S3 PUT/GET requests      & \$0.005/1k  & \$0.05  & daily pipeline + 3-hourly inference \\
Data transfer (outbound) & First 100~GB free & \$0.00 & dashboard traffic within free tier \\
GitHub Actions           & 2,000~min/month free & \$0.00 & public repo = unlimited \\
AWS Lambda               & 1M requests/month free & \$0.00 & ingestion trigger within free tier \\
\midrule
\textbf{Total (on-demand)}  & & \textbf{\textasciitilde\$15} & \\
\textbf{Total (Learner Lab)} & & \textbf{\textless\$1} & EC2 covered by credits \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Key architectural decision:} FastAPI, Streamlit, and MLflow all run on a
single shared EC2~t3.small instance. Co-location is feasible because each service
is pinned to a single worker (\texttt{uvicorn --workers 1}, \texttt{gunicorn -w 1}
for MLflow) and the 2~GB swap file absorbs transient memory spikes. Evidently runs
as an on-demand CLI invoked from GitHub Actions, not as a persistent web service ---
so the always-on surface is effectively three processes (FastAPI $\approx$ 350~MB,
MLflow UI $\approx$ 80~MB, Streamlit $\approx$ 220~MB), well within the
2~GB RAM envelope under a single-station, low-traffic workload.

\textbf{Cost optimisation options:}
\begin{itemize}
    \item \textbf{EC2 Spot Instance:} t3.small spot price is typically \$0.006/hr
          --- a 70\% reduction. Acceptable for inference workloads that can tolerate
          interruption with a restart policy.
    \item \textbf{Stop EC2 when idle:} For a capstone demonstration, stopping the
          instance between presentations reduces cost to near zero.
    \item \textbf{Scheduled shutdown:} EventBridge rules can stop the instance
          overnight (21:00--06:00) and on weekends, cutting always-on cost by
          $\approx$45\% without architectural change.
\end{itemize}

At \textasciitilde\$15/month on-demand (or under \$1/month with AWS Learner Lab
credits), CAAS is substantially cheaper than commercial air quality services
(e.g.\ IQAir Pro at \$299/year $\approx$ \$25/month for a single station),
while providing superior multi-day forecasting capability.
