\chapter{Conclusion and Recommendations}
\label{ch:conclusion}

This chapter consolidates the principal findings of the CAAS project,
discusses the limitations observed during its design and evaluation,
and outlines concrete recommendations and directions for future
research. It is organised into three sections: conclusions drawn from
the empirical results, recommendations for operational adoption, and
future research directions.

\section{Conclusions}

This project designed, implemented, and deployed CAAS --- a complete
end-to-end MLOps pipeline for multi-horizon PM2.5 forecasting in Chiang
Mai, Thailand. The principal empirical and engineering findings are
summarised below.

\subsection{Gradient-Boosted Trees Outperform LSTM at All Horizons}

Across t+1, t+3, and t+7 forecast horizons, gradient-boosted tree
ensembles outperformed the tuned LSTM comparison model. The LightGBM
champion achieves MAE of 5.12, 6.77, and 8.25~\textmu g/m\textsuperscript{3}
at t+1, t+3, and t+7 respectively, edging the XGBoost strong-secondary
(5.31 / 7.08 / 8.90) while both remain interpretable via SHAP and train
in minutes on commodity hardware. The LSTM provides competitive results
only at t+7 (MAE 8.06) but degrades substantially at t+3
(R\textsuperscript{2}~=~0.54 versus 0.73 for the gradient-boosted
models). This empirical result answers RQ2: tabular gradient boosting
with engineered lag features is the superior approach for this problem,
and XGBoost is therefore retained as a runtime fallback on the FastAPI
service to support A/B comparison, not because it wins on accuracy.

\subsection{NASA FIRMS Fire Data Adds Measurable Value}

The Scenario~D ablation study shows that removing the six FIRMS-derived
features increases MAE by 4.3--6.8\% across horizons. The effect is
strongest at t+1, where recent hotspot counts directly predict
next-day PM2.5 accumulation. This answers RQ3: fire-hotspot data is
not merely a statistically significant feature, it is operationally
material, and the cost of the FIRMS data integration is justified.

\subsection{The Predictive Signal Shifts With Horizon}

SHAP interpretation on the gradient-boosted family reveals a clear
transition in dominant features: at t+1, recent PM2.5 (\texttt{pm25\_lag1})
dominates; at t+3, 14-day fire accumulation
(\texttt{hotspot\_14d\_roll}) is most important; at t+7, seasonal
position (\texttt{sin\_month}) is the strongest signal. This gradient
--- from measurement-driven to climatology-driven prediction --- is
physically interpretable and consistent with the meteorology of Chiang
Mai haze events.

\subsection{The Default 50~\textmu g/m\textsuperscript{3} Alert Threshold Is Near-Optimal}

Scenario~C threshold tuning shows that only the t+1 classifier benefits
meaningfully from threshold adjustment (53.7~\textmu g/m\textsuperscript{3},
+4.0\% F1 gain). The standard WHO/Thai threshold is well-calibrated for
the t+3 and t+7 horizons, so the default 50~\textmu g/m\textsuperscript{3}
threshold is retained operationally.

\subsection{The MLOps Pipeline Is Production-Ready Within Its Design Scope}

Automated daily ingestion, drift monitoring with a seasonal-aware
retrain policy, champion/challenger validation, and Terraform-provisioned
AWS infrastructure together constitute a deployable, maintainable
system operating at approximately \$0.55/day. This answers RQ4:
a complete, self-retraining environmental MLOps system can be operated
within a student-scale budget.

\section{Recommendations}

The following recommendations are directed at stakeholders who wish to
adopt or extend CAAS in an operational or educational setting.

\subsection{For Operational Adoption}

\begin{itemize}
    \item \textbf{Retain both gradient-boosted models in production.}
          The LightGBM champion is the primary serving model, but the
          XGBoost strong-secondary is low-cost to keep loaded and
          provides a natural A/B comparison surface when new champions
          are promoted.

    \item \textbf{Keep the default 50~\textmu g/m\textsuperscript{3}
          alert threshold.} The threshold-tuning analysis shows that
          the aligned WHO/Thai national standard is near-optimal for
          two of the three horizons, and operational consistency with
          public-health guidance outweighs the marginal F1 gain at t+1.

    \item \textbf{Operate with the seasonal-aware drift policy.}
          Naive PSI/KS drift alarms fire during the annual haze season
          by design of the underlying data; the two-tier core/soft
          feature policy described in Chapter~\ref{ch:mlops} preserves
          genuine drift sensitivity while suppressing expected seasonal
          variation.

    \item \textbf{Shut down the always-on EC2 instance between runs
          outside demonstration periods.} Scheduled EventBridge stops
          reduce operational cost by approximately 45\% without
          architectural change.
\end{itemize}

\subsection{For Educational Adoption}

\begin{itemize}
    \item The appendix reproduction guide is sufficient to rebuild the
          system from a clean checkout in a single afternoon on a
          modern laptop, modulo LSTM training time. This makes CAAS a
          suitable reference implementation for subsequent capstone
          cohorts studying environmental MLOps.

    \item All external data sources are free at research scale (PCD
          Thailand, Open-Meteo, NASA FIRMS), and the full AWS stack can
          be operated inside an AWS Learner Lab credit budget for
          extended periods. This removes two common barriers to
          student-scale MLOps reproduction.
\end{itemize}

\section{Future Research Directions}

Several natural extensions remain outside the scope of the present
capstone but are well-motivated by the observed results.

\subsection{Multi-Station Expansion}

CAAS currently operates on a single Chiang Mai monitoring station.
PM2.5 varies spatially, and a network of stations would provide better
population coverage. Extending to all 60-plus PCD monitoring stations
in Thailand would enable national-scale air-quality forecasting and
would provide the training scale at which graph neural networks
(GNNs) become competitive for modelling spatial dependencies between
nearby stations.

\subsection{Weather Forecast Integration}

The pipeline currently uses ERA5 reanalysis weather data, which is
accurate but historical. True production use at t+3 and t+7 horizons
requires weather-forecast inputs (ECMWF HRES or GFS) as upstream
features. Incorporating forecast meteorology would introduce forecast
uncertainty into the pipeline but is essential for genuine multi-day
predictive operation rather than simulated hindcast evaluation.

\subsection{Improved Long-Horizon Modelling}

Despite architectural improvements (sequence length~=~60, deeper
network, batch normalisation), the LSTM R\textsuperscript{2} of 0.54 at
t+3 indicates the model struggles at the mid-range horizon. This
suggests the temporal sequences do not provide sufficient additional
signal over the tabular lag features already engineered for the
gradient-boosted models. Transformer architectures with learned
temporal positional encodings, or explicit separation of climatological
and anomaly components, are promising directions for closing this
gap at t+3.

\subsection{Seasonal-Component Modelling for Sparse-Fire Periods}

Zero-inflation in the FIRMS features during the rainy season limits
PSI-based drift detection and requires suppression logic.
Explicit seasonal-component models --- for example, STL decomposition
with separate training regimes for the burning and non-burning
seasons --- could handle this more elegantly than the current
suppression heuristic.

\subsection{Kubernetes Serving at Scale}

For production deployment at scale, migrating from a single EC2
instance to a Kubernetes cluster with horizontal pod autoscaling
would remove the co-location ceiling on request throughput and enable
geographically distributed serving. This is unnecessary at the current
single-station, low-traffic workload but is the obvious scaling path
beyond it.

\subsection{Deeper Hyperparameter Optimisation}

The current Optuna searches (100 trials per family per horizon) could
be extended with multi-fidelity pruning (Hyperband) or
population-based training for more efficient use of compute, and the
search space itself could be broadened to include regularisation
schedules and early-stopping variants that the current search holds
fixed.

\subsection{Limitations}

Finally, we acknowledge the limitations of the present work, which
bound the generalisability of our conclusions. CAAS operates on a
single Chiang Mai monitoring station; PM2.5 varies spatially and
inter-station transferability is not demonstrated. The pipeline uses
historical ERA5 weather (accurate but retrospective) rather than
forecast weather inputs, so the t+3 and t+7 evaluations are hindcasts
rather than genuine forward predictions. The LSTM t+3
underperformance was not fully resolved within the project budget.
Zero-inflation in the FIRMS features during the rainy season limits
PSI-based drift detection and currently requires suppression logic
rather than a more principled seasonal model. Addressing these
limitations constitutes the core of the future-work agenda above.
