\chapter*{References}
\addcontentsline{toc}{chapter}{References}

\begin{thebibliography}{99}

% --- Data Sources ---

\bibitem{pcd}
Pollution Control Department (PCD), Thailand. \textit{Air4Thai Air Quality
Reporting System}. Ministry of Natural Resources and Environment.
\url{http://air4thai.pcd.go.th/}. Accessed 2025.

\bibitem{openmeteo}
Zippenfenig, P. (2023). Open-Meteo.com Weather API.
Zenodo. \url{https://doi.org/10.5281/zenodo.7970649}.
\url{https://open-meteo.com/}

\bibitem{firms}
NASA FIRMS. \textit{Fire Information for Resource Management System —
VIIRS/SNPP Active Fire Product}. NASA Earth Science Data and Information
System (ESDIS). \url{https://firms.modaps.eosdis.nasa.gov/}. Accessed 2025.

\bibitem{era5}
Hersbach, H. et al. (2020). The ERA5 global reanalysis.
\textit{Quarterly Journal of the Royal Meteorological Society},
146(730), 1999--2049. \url{https://doi.org/10.1002/qj.3803}

% --- Machine Learning ---

\bibitem{xgboost}
Chen, T. and Guestrin, C. (2016). XGBoost: A Scalable Tree Boosting System.
\textit{Proceedings of the 22nd ACM SIGKDD International Conference on
Knowledge Discovery and Data Mining (KDD '16)}, 785--794.
\url{https://doi.org/10.1145/2939672.2939785}

\bibitem{shap}
Lundberg, S.~M. and Lee, S.-I. (2017). A Unified Approach to Interpreting
Model Predictions. \textit{Advances in Neural Information Processing Systems
(NeurIPS)}, 30. \url{https://arxiv.org/abs/1705.07874}

\bibitem{lstm}
Hochreiter, S. and Schmidhuber, J. (1997). Long Short-Term Memory.
\textit{Neural Computation}, 9(8), 1735--1780.
\url{https://doi.org/10.1162/neco.1997.9.8.1735}

\bibitem{huber}
Huber, P.~J. (1964). Robust estimation of a location parameter.
\textit{The Annals of Mathematical Statistics}, 35(1), 73--101.

\bibitem{sklearn}
Pedregosa, F. et al. (2011). Scikit-learn: Machine Learning in Python.
\textit{Journal of Machine Learning Research}, 12, 2825--2830.
\url{https://jmlr.org/papers/v12/pedregosa11a.html}

\bibitem{keras}
Chollet, F. et al. (2015). Keras. \url{https://keras.io/}

% --- MLOps and Infrastructure ---

\bibitem{mlflow}
Zaharia, M. et al. (2018). Accelerating the Machine Learning Lifecycle with
MLflow. \textit{IEEE Data Engineering Bulletin}, 41(4), 39--45.
\url{https://databricks.com/wp-content/uploads/2018/12/MLflow-Dec18.pdf}

\bibitem{evidently}
Evidently AI. \textit{Evidently: Open-Source ML Monitoring}.
\url{https://www.evidentlyai.com/}. Accessed 2025.

\bibitem{fastapi}
Ramírez, S. (2019). FastAPI. \url{https://fastapi.tiangolo.com/}.
Accessed 2025.

\bibitem{terraform}
HashiCorp. \textit{Terraform: Infrastructure as Code}.
\url{https://www.terraform.io/}. Accessed 2025.

\bibitem{docker}
Merkel, D. (2014). Docker: Lightweight Linux Containers for Consistent
Development and Deployment. \textit{Linux Journal}, 2014(239).

\bibitem{githubactions}
GitHub. \textit{GitHub Actions: Automate Your Workflow}.
\url{https://docs.github.com/en/actions}. Accessed 2025.

% --- Domain / Air Quality ---

\bibitem{who2021}
World Health Organization (2021). \textit{WHO Global Air Quality Guidelines:
Particulate Matter (PM2.5 and PM10), Ozone, Nitrogen Dioxide, Sulfur
Dioxide and Carbon Monoxide}. WHO Press, Geneva.
\url{https://www.who.int/publications/i/item/9789240034228}

\bibitem{psi}
Yap, X.~Q. and Hashim, M. (2013). A robust calibration approach for PM10
prediction from MODIS aerosol optical depth. \textit{Atmospheric Chemistry
and Physics}, 13(6), 3517--3526.

\bibitem{ks_test}
Kolmogorov, A.~N. (1933). Sulla determinazione empirica di una legge di
distribuzione. \textit{Giornale dell'Istituto Italiano degli Attuari}, 4, 83--91.

\bibitem{chaing_mai_haze}
Bontemps, C. et al. (2021). Agricultural burning and PM2.5 in Southeast Asia:
regional model assessment. \textit{Atmospheric Environment}, 248, 118218.

\end{thebibliography}
